Stellar Bars in Isolated Gas-Rich Galaxies Do Not Slow Down

### 2022-04-08 paper/ms.tex
Add bar analysis section

>>>>

\noindent
{\bf Bar analysis}
\\
\noindent
The analysis of various bar properties is performed as follows. First, the
pattern speed is measured from the angle of the second Fourier component. We measured the second Fourier component by computing,
\begin{equation}
\begin{split}
A_2 &= \sum_i m_i e^{i 2 \phi_i} \\
A_0 &= \sum_i m_i \textrm{,}
\end{split}
\end{equation}
where $m_i$ and $\phi_i$ are the mass and azimuthal angle of each particle, respectively. We
computed $A_2$ and $A_0$ in cylindrical bins of width $0.5\,\textrm{kpc}$ from radii of
$0$ to $30\,\textrm{kpc}$. We defined the angle of the bar $\phi_b$ to be
twice the angle of complex number $A_2$ as measured in the bin extending from
a radius of $2.5$ to $3\,\textrm{kpc}$. After correcting for the periodicity
of $\phi_b$, we measured the pattern speed as the two-sided finite gradient of
$\phi_b$ as a function of time.

The time evolution of the bar strength, defined as the maximum of
$\left|A_2/A_0\right|$ as a function of radius, is shown in Extended Data
Fig.~\ref{fig:strength}. The quantity $\left|A_2/A_0\right|$ varies from $0$
to $1$, with larger values indicating a stronger bar pattern. We see that in
the N-body case, $\left|A_2/A_0\right|$ increases over time as the bar pattern
slows. This is consistent with previous N-body simulations which showed a
clear correlation between the bar pattern speed and the bar
strength.\cite{2003MNRAS.341.1179A} In the SMUGGLE case, we see that the bar
strength has an initial drop but then remains at a roughly constant but
slightly decreasing strength. This is consistent with the pattern speed in the
SMUGGLE case being roughly constant or slightly increasing.

\begin{figure}[h]%
\centering
\includegraphics[width=0.45\textwidth]{fig/fig-A2.pdf}
\caption{The time evolution of the bar strength, measured as the second
Fourier component divided by the zeroth Fourier component (a formula is given
in the Methods section). We see that in the N-body case (blue) the bar
strength increases with time, consistent with previous results showing that
the bar strength increases as bars slow down. In the SMUGGLE case (orange), we
see that the bar strength remains roughly constant, possibly slightly
decreasing with time. This is also consistent with the expected relation
between pattern speed and strength since the bar in this case is not slowing
down.}
\label{fig:strength}
\end{figure}

Computing the length of the bar and the torque on the bar by different
components requires us to decompose the disk into a component which is trapped
by the bar and a component which is untrapped. In order to do this, we follow
closely the technique developed in ref.\cite{2016MNRAS.463.1952P} We analyzed
the orbit of each star particle (meaning initial disk, bulge, and newly formed
stars) by extracting the $x$-$y$ positions of the apoapse of each in a frame
corotating with the bar (apoapses are defined as local maxima in $r$). For
each apoapse, we searched for the $19$ closest apoapses in time and applied a
$k$-means clustering algorithm on this set of $20$ points with $k=2$. We then
computed for each of the two clusters the average angle from the bar
$\left<\Delta \phi\right>_{0,1}$, the standard deviation in $R$ of the points
${\sigma_R}_{0,1}$, and the average radius of the cluster
$\left<R\right>_{0,1}$. At each apoapse, a particle was considered to be in
the bar if it met the following criteria:
\begin{equation}
\textrm{max}\left(\left<\Delta \phi\right>_{0,1}\right) < \pi / 8
\end{equation}
\begin{equation}
\frac{{\sigma_R}_0 + {\sigma_R}_1}{\left<R\right>_0 + \left<R\right>_1} < 0.22
\end{equation}
These criterion are slightly different and simplified from the ones used in
ref.\cite{2016MNRAS.463.1952P}, but we found to empirically work well at
decomposing the disk into a bar and disk component. In Extended Data
Fig.~\ref{fig:decomp}, we show an example of this decomposition. The
\textit{left} panel shows a surface density projection of the stellar disk and
bulge (including newly formed stars) from the SMUGGLE model after
$1\,\text{Gyr}$ of evolution in a frame such that the bar is aligned with the
$x$-axis. The \textit{middle} panel shows a projection of the subset of stars
that are identified as being trapped in the bar and the \textit{right} panel
shows a projection of the stars that are not identified as being trapped. The
fact that the \textit{right} panel is roughly axisymmetric indicates the bar
decomposition is performing adequately.

\begin{figure*}[h]%
\centering
\includegraphics[width=0.95\textwidth]{fig/fig-bar_decomp.pdf}
\caption{Demonstration of our bar decomposition procedure based on
ref.\cite{2016MNRAS.463.1952P} The \textit{left panel} shows a surface density
projection through the stellar component of the N-body simulation (disk and
bulge). The \textit{middle panel} shows the component of the disk identified
as being trapped in the bar while the \textit{right panel} shows the component
of the disk identified as not being trapped in the bar. The fact that the
untrapped stars form a roughly axisymmetric structure indicates our bar
decomposition is sufficiently accurate. {\color{red} I will remake this figure
- I ended up losing the files which stored the bar membership, so I just
pulled these screenshots from an earlier presentation.}}
\label{fig:decomp}
\end{figure*}

After the disk has been decomposed into a trapped and untrapped component, we
measured the bar length as being the radius $R_b$ which encapsulates $99\%$ of
the stars identified as being trapped in the bar, allowing for some outliers.
For the computation of the torque, we used the tree algorithm in
\texttt{MakeNewDisk} customized to be accessible from \texttt{Python} using
\texttt{Cython}. This algorithm is based on the \texttt{TREESPH} code. We
constructed a tree using only the star particles identified as being trapped
in the bar. We then queried the tree at the locations of all resolution
elements in the other components and computed the torque of the bar on such
components. The torque on the bar by the other components is simply the
negative of the torque on the other components by the bar. A similar method
was applied in measuring the torque for the disk when the pattern speed is
kept constant, as is done in Fig.~\ref{fig:equil}.
